---
title: "01a ICB - Prep Domain Numerator"
author: "J Andres Gannon"
date: "May 19, 2020"
output:
  html_document:
    theme: flatly
    code_download: true
    df_print: paged
    number_sections: yes
    toc: yes
    toc_depth: '5'
    toc_float:
      collapsed: yes
editor_options:
  chunk_output_type: inline
---

```{r setup, include = FALSE}
library(magrittr)
library(ggplot2)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, echo = FALSE)
```

This document loads in the newly created ICB data that identifies the military domains used by each crisis-actor.

# Load
Start by loading the cleaned ICB data. For this version of the data, each row is a crisis-actor and each column is a dummy variable for a military domain
```{r}
df <- readRDS(file = paste0(here::here(), '/data/00_crisis_actor_domain_type.Rds'))
```

```{r, eval = FALSE}
# REPLICATION NOTE: the accepted paper version was built using a beta version of the ICBe data that is manually edited and re-coded here. Scholars replicating this piece should use ICBe_V1_wide_agreed (https://github.com/CenterForPeaceAndSecurityStudies/ICBEdataset/blob/master/replication_data/out/ICBe_V1_wide_agreed.Rds) for different codings of military domains that are described in the accompanying paper
```

We can examine distributions of the numeric variables to identify outliers or problems in data quality. Everything looks largely correct
```{r, fig.width = 16, fig.height = 16}
DT::datatable(funModeling::df_status(df))
funModeling::plot_num(df)
```

# Prep ICB domain data
## Reshape
We need the ICB domains data with each row being a crisis actor and each column is a dummy for type-domain

Each domain can be "used" by an actor as either a thought, speech, action. We want to disaggregate those because they are not the same. That can also tell us about escalation patterns. We use the same process to convert the wikipedia URL actor names into strings that can be matched to GW country codes and qcodes
```{r}
# Convert event_type column to lower case
df$event_type <- tolower(df$event_type)
```

We then convert the event_type column into the appropriate dummies
```{r}
# Make blank columns for each domain_type
df$land_action <- NA
df$land_speech <- NA
df$land_thought <- NA

df$air_action <- NA
df$air_speech <- NA
df$air_thought <- NA

df$sea_action <- NA
df$sea_speech <- NA
df$sea_thought <- NA

df$wmd_action <- NA
df$wmd_speech <- NA
df$wmd_thought <- NA

df$space_action <- NA
df$space_speech <- NA
df$space_thought <- NA

df$cyber_action <- NA
df$cyber_speech <- NA
df$cyber_thought <- NA

# If event_type = X and domain = 1, domain_type = 1. Otherwise, 0
df$land_action[df$event_type == "action" &
                      df$land == 1] <- 1
df$land_speech[df$event_type == "speech" &
                      df$land == 1] <- 1
df$land_thought[df$event_type == "thought" &
                      df$land == 1] <- 1

df$air_action[df$event_type == "action" &
                      df$air == 1] <- 1
df$air_speech[df$event_type == "speech" &
                      df$air == 1] <- 1
df$air_thought[df$event_type == "thought" &
                      df$air == 1] <- 1

df$sea_action[df$event_type == "action" &
                      df$sea == 1] <- 1
df$sea_speech[df$event_type == "speech" &
                      df$sea == 1] <- 1
df$sea_thought[df$event_type == "thought" &
                      df$sea == 1] <- 1

df$wmd_action[df$event_type == "action" &
                      df$wmd == 1] <- 1
df$wmd_speech[df$event_type == "speech" &
                      df$wmd == 1] <- 1
df$wmd_thought[df$event_type == "thought" &
                      df$wmd == 1] <- 1

df$space_action[df$event_type == "action" &
                      df$space == 1] <- 1
df$space_speech[df$event_type == "speech" &
                      df$space == 1] <- 1
df$space_thought[df$event_type == "thought" &
                      df$space == 1] <- 1

df$cyber_action[df$event_type == "action" &
                      df$cyber == 1] <- 1
df$cyber_speech[df$event_type == "speech" &
                      df$cyber == 1] <- 1
df$cyber_thought[df$event_type == "thought" &
                      df$cyber == 1] <- 1

# Drop event type and group by crisis actor
df <- df %>% 
  dplyr::select(-c(event_type, land, air, sea, wmd, space, cyber)) %>%
  dplyr::filter(!is.na(actor)) %>%
  dplyr::group_by(crisno, actor) %>%
  replace(is.na(.), 0) %>%
  dplyr::summarise_all(sum)
```

## Add year and crisis name
These variables are helpful for getting the correct actor codes, since they change over time, and just generally identifying the crises
```{r}
# Load
icb <- read.csv(file = paste0(here::here(), '/inst/extdata/icb1v13.csv'))

# Subset to variables of interest and rename
icb <- icb %>% dplyr::select(crisno, crisname, yrtrig) %>%
  dplyr::rename(year = yrtrig)

# Merge
df <- dplyr::left_join(df, icb)
```

## Fix actor names
Actor names are currently coded using their wikipedia URL. We want to convert that to the corresponding wikidata names with Q codes
```{r}
# Ungroup since it's causing a lot of problems
df <- df %>% dplyr::ungroup()

# Keep only actor name
df$actor <- sub(".*/", "", df$actor)

# Remove underscores
df$actor <- stringr::str_replace_all(df$actor, "_", " ")

# Clean UTF-8 encoding errors
df$actor[df$actor == "%C3%85land Islands"] <- "Åland Islands"
df$actor[df$actor == "%C3%89tienne Tshisekedi"] <- "Étienne Tshisekedi"
df$actor[df$actor == "%C3%91ancahuaz%C3%BA Guerrilla"] <- "Ñancahuazú Guerrilla"
df$actor[df$actor == "All-African Peoples%27 Conference"] <- "All-African Peoples' Conference"
df$actor[df$actor == "King%27s African Rifles"] <- "King's African Rifles"
df$actor[df$actor == "Mongolian People%27s Republic"] <- "Mongolian People's Republic"
df$actor[df$actor == "People%27s Armed Forces"] <- "People's Armed Forces"

# Set to lower case for easier merging
df$actor <- tolower(df$actor)
```

# Add external variables
## Qcodes
We now add qcodes for each actor which is their unique identifier from wikidata. These have been manually collected for each domain crisis-actor. The table shown below are all the non-Gleditsch Ward country crisis-actors for which we have domain data. The table should represent international bodies and non-state actors
```{r}
qcodes <- read.csv(file = paste0(here::here(), '/data/icb_actor_qcodes.csv'))

# Prep Qcodes for merging
qcodes <- qcodes %>% 
  dplyr::select(actor, qcode, trash_actor, gwcode, gwname) %>%
  dplyr::mutate(actor = tolower(actor))

# Merge manual Qcodes into the ICB domain dataframe
df <- dplyr::left_join(df, qcodes)

# Remove trash actors
df <- df %>% 
  dplyr::filter(is.na(trash_actor)) %>%
  dplyr::select(-trash_actor)

# Inspect
df %>% 
  dplyr::select(actor, qcode, gwname, gwcode) %>%
  dplyr::filter(is.na(gwcode)) %>%
  dplyr::select(-gwcode, -gwname) %>%
  dplyr::distinct() %>%
  dplyr::arrange(actor) %>%
  DT::datatable()
```

## GW country codes
While in theory the Qcodes should be matched to GW country codes in the previous dataset, they are not matched by year which produces problems for entities that change code over time or should only be considered states during certain time periods
```{r}
df_mssl <- read.csv(file = paste0(here::here(), '/data/polities_mssl_gleditschward_states.csv'))

# Subset to columns of interest
df_mssl <- df_mssl %>% 
  dplyr::select(nation_wikidata_id, gwcode, gwabrev, gw_start, gw_end, microstate) %>%
  dplyr::filter(microstate == 0) %>%
  dplyr::rename(qcode = nation_wikidata_id, 
                gwid_mssl = gwcode, 
                gwname_mssl = gwabrev) %>%
  dplyr::select(-microstate)

# To capture post-2021 crises, change end date for those ending in 2012 to be 2020
df_mssl$gw_end[df_mssl$gw_end == "2012-12-31"] <- "2020-12-31"

# Make long to expand year range, since we need to match on the qcode-year combo
df_mssl$gw_start <- as.POSIXct(df_mssl$gw_start)
df_mssl$gw_end <- as.POSIXct(df_mssl$gw_end)

df_mssl$gw_start <- as.numeric(format(df_mssl$gw_start, '%Y'))
df_mssl$gw_end <- as.numeric(format(df_mssl$gw_end, '%Y'))

# Make long
df_mssl <- df_mssl %>% dplyr::mutate(year = purrr::map2(gw_start, gw_end, `:`)) %>%
  dplyr::select(-gw_start, -gw_end) %>%
  tidyr::unnest(cols = year) %>%
  dplyr::rename(gwcode_date = gwid_mssl,
                gwname_date = gwname_mssl)

# Merge
df <- dplyr::left_join(df, df_mssl)

# Identify the entities that have different GW codes
df %>%
  dplyr::mutate(diff = dplyr::if_else(gwcode == gwcode_date, 0, 1)) %>%
  dplyr::filter(diff == 1) %>%
  dplyr::select(crisno, year, actor, qcode, dplyr::starts_with("gw")) %>%
  dplyr::arrange(actor) %>%
  dplyr::filter(stringr::str_detect(actor, 'vietnam')) %>%
  DT::datatable()

# Fix date-based discrepancies
## Czech Republic
df$gwcode[df$gwcode == 316 &
            df$year < 1993] <- 315
df$gwname[df$gwname == "CZR" &
            df$year < 1993] <- "CZE"

## Yugoslavia
df$gwcode[df$gwcode == 340 &
            df$year < 2006] <- 345
df$gwname[df$gwname == "SER" &
            df$year < 2006] <- "YUG"

## Vietnam
df$gwcode[df$gwcode == 815 &
            df$year == (1953 | 1977)] <- 817
df$gwname[df$gwname == "VNM" &
            df$year == (1953)] <- "RVN"

df$gwcode[df$gwcode == 815 &
            df$year > 1974] <- 816
df$gwname[df$gwname == "VNM" &
            df$year > 1974] <- "DRV"

## Germany
df$gwcode[df$gwcode == 260 &
            df$year < 1950] <- 255
df$gwcode[df$gwcode == 265 &
            df$year < 1950] <- 255
df$gwname[df$gwcode == 255] <- "GMY"
```

Now that that's cleaned, the list of crisis-actors without a GW code should only refer to non-state or suprastate actors
```{r}
df %>%
  dplyr::filter(is.na(gwcode)) %>%
  dplyr::select(actor, gwname, gwcode) %>%
  dplyr::distinct() %>%
  dplyr::arrange(actor) %>%
  DT::datatable()
```

# Sort
We re-sort the columns and make sure they are appropriately named
```{r}
df <- df %>% 
  dplyr::rename(gwid = gwcode) %>%
  dplyr::select(crisno, crisname, year,
                actor, gwname, gwid,
                land_action, air_action, sea_action, wmd_action, space_action, cyber_action)

funModeling::plot_num(df)
```

# Fix edge cases
## Actor coding
Some actors still have GW codes even though they do not refer to the country. We fix those manually
```{r}
df %>%
  dplyr::select(actor, gwname, gwid) %>%
  dplyr::distinct() %>%
  dplyr::arrange(gwname) %>%
  DT::datatable()

df$gwname[df$actor == "national congress (sudan)"] <- NA
df$gwid[df$actor == "national congress (sudan)"] <- NA

df$gwname[df$actor == "syrian opposition"] <- NA
df$gwid[df$actor == "syrian opposition"] <- NA

df$gwname[df$actor == "trilateral contact group on ukraine"] <- NA
df$gwid[df$actor == "trilateral contact group on ukraine"] <- NA
```

## Domain coding
Some action codings have to be fixed. We double check all action codings and manually fix ones that were errors by the original coders. We start by examining the less common domains of WMD, space, and cyber
```{r}
# Land
## Cases coded 1 that should be 0
  ### US during north korea satellite launch 2009
  df$land_action[df$crisno == 459 &
                    df$gwname == "USA"] <- 0

  ### US during north korea satellite launch 2009
  df$land_action[df$crisno == 459 &
                    df$gwname == "ROK"] <- 0

  ### Japan during north korea satellite launch 2009
  df$land_action[df$crisno == 459 &
                    df$gwname == "JPN"] <- 0

# Sea
## Cases coded 1 that should be 0
  ### North Korea during north korea satellite launch 2009
  df$sea_action[df$crisno == 459 &
                    df$gwname == "PRK"] <- 0
  
# WMD
## Cases coded 1 that should be 0
  ### Bolivia during chaco 1 1928
  df$wmd_action[df$crisno == 37 &
                    df$gwname == "BOL"] <- 0

  ### Japan during Shanghai crisis 1932
  df$wmd_action[df$crisno == 40 &
                    df$gwname == "JPN"] <- 0

  ### Ethiopia during Ethiopian war 1934
  df$wmd_action[df$crisno == 47 &
                    df$gwname == "ETH"] <- 0

  ### Italy during Ethiopian war 1934
  df$wmd_action[df$crisno == 47 &
                    df$gwname == "ITA"] <- 0

  ### UK during Ethiopian war 1934
  df$wmd_action[df$crisno == 47 &
                    df$gwname == "UKG"] <- 0

  ### Russia during Fall of Saipan 1944
  df$wmd_action[df$crisno == 95 &
                    df$gwname == "RUS"] <- 0

  ### Egypt during Yum Kippus war 1973
  df$wmd_action[df$crisno == 255 &
                    df$gwname == "EGY"] <- 0
    
  ### France during Iraq nuclear reactor 1981
  df$wmd_action[df$crisno == 324 &
                    df$gwname == "FRN"] <- 0

  ### China during Stratleys 1988
  df$wmd_action[df$crisno == 384 &
                    df$gwname == "CHN"] <- 0

  ### China during Taiwan Strait 1996
  df$wmd_action[df$crisno == 415 &
                    df$gwname == "CHN"] <- 0

  ### Syria during Haifa suicide bombing 2003
  df$wmd_action[df$crisno == 443 &
                    df$gwname == "SYR"] <- 0
  
  ### Japan during NK nuclear 2006
  df$wmd_action[df$crisno == 450 &
                    df$gwname == "JPN"] <- 0
  
  ### South Korea during NK nuclear 2006
  df$wmd_action[df$crisno == 450 &
                    df$gwname == "ROK"] <- 0
  
  ### US during NK nuclear 2013
  df$wmd_action[df$crisno == 469 &
                    df$gwname == "USA"] <- 0

## Cases coded 0 that should be 1
  ### Egypt during Yemen war III 1965
  df$wmd_action[df$crisno == 212 &
                    df$gwname == "EGY"] <- 1

  ### US during Tet offensive 1968
  df$wmd_action[df$crisno == 225 &
                    df$gwname == "USA"] <- 1

  ### Rhodesia during Rhodesia raid 1977
  df$wmd_action[df$crisno == 283 &
                    df$gwname == "ZIM"] <- 1

  ### Iraq during Iraw recapture of Al-Faw 1988
  df$wmd_action[df$crisno == 385 &
                    df$gwname == "IRQ"] <- 1

  ### Syria during Syria chemical weapons 2013
  df$wmd_action[df$crisno == 470 &
                    df$gwname == "SYR"] <- 1

# Space
## Cases coded 1 that should be 0
  ### USA during Russian civil war 1918
  df$space_action[df$crisno == 3 &
                    df$gwname == "USA"] <- 0
  
  ### Turkey during Greece-Turkey war 1922
  df$space_action[df$crisno == 25 &
                    df$gwname == "TUR"] <- 0
  
  ### Japan during Shanghai crisis 1932
  df$space_action[df$crisno == 40 &
                    df$gwname == "JPN"] <- 0
  
  ### USA during Cuban Missile Crisis 1962
  df$space_action[df$crisno == 196 &
                    df$gwname == "USA"] <- 0
  
  ### Pakistan during I-P nuclear tests 1998
  df$space_action[df$crisno == 425 &
                    df$gwname == "PAK"] <- 0
  
  ### US during North Korea nuclear test 2013
  df$space_action[df$crisno == 469 &
                    df$gwname == "USA"] <- 0

## Cases coded 0 that should be 1
  ### North Korea during North Korea nuclear test satellite launch 2009
  df$space_action[df$crisno == 459 &
                    df$gwname == "PRK"] <- 1

# Cyber
## Cases coded 1 that should be 0
  ### Japan during Shanghai 1932
  df$cyber_action[df$crisno == 40 &
                    df$gwname == "JPN"] <- 0

  ### Egypt during Formation of UAR 1958
  df$cyber_action[df$crisno == 162 &
                    df$gwname == "EGY"] <- 0

  ### Syria during formation of UAR 1958
  df$cyber_action[df$crisno == 162 &
                    df$gwname == "SYR"] <- 0

## Cases coded 0 that should be 1
  ### Russia cyber attack during Turkey Russia jet incident 2016
  df$cyber_action[df$crisno == 476 &
                    df$gwname == "RUS"] <- 1
```

# Save
Save final version of the new data
```{r}
saveRDS(df, paste0(here::here(), "/data/","01a_icb_domains_monadic.rds"))
```
