---
title: '\singlespacing One if by land, and two if by sea: Cross-domain contests and the escalation of international crises'
author: |
  | J Andr$\'{e}$s Gannon^[Email: \texttt{jagannon@ucsd.edu}. Web: \texttt{jandresgannon.com}. \newline The author wishes to thank Erik Gartzke and Jon Lindsay who were Principal Investigators on Office of Naval Research Grant N00014-14-1-0071 and the Department of Defense Minerva Research Initiative grant that organized the data collection for this project. Rex Douglass and Thomas Leo Scherer are responsible for the underlying ontology and event data and ICB project managers Jonathan Wilkenfeld and David Quinn oversaw the coding effort. Tom Brailey, David McCrum, James McCullough, Cole Reynolds, Cailen Rodriguez, Woong Suh, and Erin Werner provided excellent research assistance. The author is also grateful for constructive feedback from Shannon Carcelli, Austin Carson, Jasen Castillo, James Fowler, Patrick Hulme, Igor Kovak, Kendrick Kuo, Brian Lai, David Lake, Nicole Letourneau, Saadia Pekkanen, Sara Plana, Laura Resnick Samotin, Branislav Slantchev, Rachel Schoner, Sanne Verschuren, Barbara Walter, and workshop participants at UC San Diego and the Charles Koch Foundation working groups. Previous versions were presented at the 2017 Peace Science Society (International) conference and 2018 International Studies Association conference. Any opinions, findings, or recommendations do not necessarily reflect the view of the Office of Naval Research. Data and code for the empirical analysis can be found at \texttt{https://github.com/CenterForPeaceAndSecurityStudies/icb}.]
  | Department of Political Science
  | University of California, San Diego
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output:
  bookdown::pdf_document2:
    number_sections: yes
    toc: no
    keep_tex: yes
  bookdown::html_notebook2:
    fig_height: 8
    fig_width: 12
    number_sections: yes
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float:
      collapsed: yes
  bookdown::word_document2: default
  bookdown::word_document:
    toc: no
site: bookdown::bookdown_site
editor_options:
  chunk_output_type: inline
abstract: \singlespacing The emergence of new modes of conflict has prompted inquiry about the relationship between conflict escalation and how states fight. Despite  rich theoretical literature contrasting deterrence and spiral models of conflict more generally, empirical work on different domains of military action --- land, air, sea, WMD, space, and cyber --- has been largely constrained to single-domain studies, limiting our understanding of how these domains interact. This paper contributes to our understanding of cross-domain military interactions by introducing a new dataset of the domains in which states took military action during 412 international crises from 1918 to 2010. In doing so, it tests existing theories about cross-domain deterrence and cross-domain conflict to identify the relationship between conflict escalation and belligerents interacting with unlike means. We demonstrate that cross-domain interactions are not rare -- they are instead the modal form of conflict -- but despite contemporary concern, this phenomenon is no more common today than aggression with unlike means. Crises with cross-domain military interactions are less violent and no longer than crises in which belligerents engage with similar means. \linebreak \linebreak Word count (excluding references) `r wordcountaddin::word_count()`
header-includes:
- \usepackage{tikz}
- \usepackage{pgfplots}
- \pgfplotsset{compat=newest}
- \usetikzlibrary{plotmarks}
- \usetikzlibrary{arrows.meta}
- \usepgfplotslibrary{patchplots}
- \usepackage{grffile}
- \usepackage{caption}
- \usepackage[utf8]{inputenc}
- \usepackage[doublespacing]{setspace}
- \usepackage{float}
- \usepackage{multirow}
bibliography: icbdomains.bib
---

```{r knitr_options, echo = FALSE, warning = FALSE}
# added to indicate that the home holder is the repo folder
library(knitr)
library(magrittr)
library(ggplot2)
knitr::opts_knit$set(root.dir = normalizePath(here::here()))
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, error = FALSE, message = FALSE, fig.pos = "h!", out.extra = "", fig.align = "center")
```

```{r load data}
df_allnew <- readRDS(file = paste0(here::here(), '/data/01a_icb_domains_monadic.rds'))
df_actor <- readRDS(file = paste0(here::here(), '/data/01b_icb_domains_monadic.rds'))
df_crisis <- readRDS(file = paste0(here::here(), '/data/01b_icb_domains_crisis.rds'))
```

```{r fixed polr function}
# Fix polr functions
# We use the polr function from the MASS package to compute an ordered probit. The base version of the function in the R package contains an error that does not take the log of differences in the reposed zetas which results in an optimization error where vmmin is infinite. A fixed version of the function was created and is loaded here. For this reason, the polr function is not loaded from the MASS package, but instead from the function below. All secondary functions from the MASS package are compatible with the output of the revised function.

# file MASS/R/polr.R
# copyright (C) 1994-2008 W. N. Venables and B. D. Ripley
# Use of transformed intercepts contributed by David Firth
#
#  This program is free software; you can redistribute it and/or modify
#  it under the terms of the GNU General Public License as published by
#  the Free Software Foundation; either version 2 or 3 of the License
#  (at your option).
#
#  This program is distributed in the hope that it will be useful,
#  but WITHOUT ANY WARRANTY; without even the implied warranty of
#  MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
#  GNU General Public License for more details.
#
#  A copy of the GNU General Public License is available at
#  http://www.r-project.org/Licenses/
#
library(MASS)

polr <- function(formula, data, weights, start, ..., subset,
                 na.action, contrasts = NULL, Hess = FALSE,
                 model = TRUE,
                 method = c("logistic", "probit", "cloglog", "cauchit"))
{
    logit <- function(p) log(p/(1 - p))

    fmin <- function(beta) {
        theta <- beta[pc + 1L:q]
        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
        eta <- offset
        if (pc > 0)
            eta <- eta + drop(x %*% beta[1L:pc])
        pr <- pfun(gamm[y + 1] - eta) - pfun(gamm[y] - eta)
        if (all(pr > 0))
            -sum(wt * log(pr))
        else Inf
    }

    gmin <- function(beta)
    {
        jacobian <- function(theta) { ## dgamma by dtheta matrix
            k <- length(theta)
            etheta <- exp(theta)
            mat <- matrix(0 , k, k)
            mat[, 1] <- rep(1, k)
            for (i in 2:k) mat[i:k, i] <- etheta[i]
            mat
        }
        theta <- beta[pc + 1L:q]
        gamm <- c(-Inf, cumsum(c(theta[1L], exp(theta[-1L]))), Inf)
        eta <- offset
        if(pc > 0) eta <- eta + drop(x %*% beta[1L:pc])
        pr <- pfun(gamm[y+1] - eta) - pfun(gamm[y] - eta)
        p1 <- dfun(gamm[y+1] - eta)
        p2 <- dfun(gamm[y] - eta)
        g1 <- if(pc > 0) t(x) %*% (wt*(p1 - p2)/pr) else numeric(0)
        xx <- .polrY1*p1 - .polrY2*p2
        g2 <- - t(xx) %*% (wt/pr)
        g2 <- t(g2) %*% jacobian(theta)
        if(all(pr > 0)) c(g1, g2) else rep(NA, pc+q)
    }

    m <- match.call(expand.dots = FALSE)
    method <- match.arg(method)
    pfun <- switch(method, logistic = plogis, probit = pnorm,
                   cloglog = pgumbel, cauchit = pcauchy)
    dfun <- switch(method, logistic = dlogis, probit = dnorm,
                   cloglog = dgumbel, cauchit = dcauchy)
    if(is.matrix(eval.parent(m$data)))
        m$data <- as.data.frame(data)
    m$start <- m$Hess <- m$method <- m$model <- m$... <- NULL
    m[[1L]] <- as.name("model.frame")
    m <- eval.parent(m)
    Terms <- attr(m, "terms")
    x <- model.matrix(Terms, m, contrasts)
    xint <- match("(Intercept)", colnames(x), nomatch=0L)
    n <- nrow(x)
    pc <- ncol(x)
    cons <- attr(x, "contrasts") # will get dropped by subsetting
    if(xint > 0) {
        x <- x[, -xint, drop=FALSE]
        pc <- pc - 1
    } else warning("an intercept is needed and assumed")
    wt <- model.weights(m)
    if(!length(wt)) wt <- rep(1, n)
    offset <- model.offset(m)
    if(length(offset) <= 1) offset <- rep(0, n)
    y <- model.response(m)
    if(!is.factor(y)) stop("response must be a factor")
    lev <- levels(y)
    if(length(lev) <= 2) stop("response must have 3 or more levels")
    y <- unclass(y)
    q <- length(lev) - 1
    Y <- matrix(0, n, q)
    .polrY1 <- col(Y) == y
    .polrY2 <- col(Y) == y - 1
    if(missing(start)) {
      # try something that should always work -tjb
      u <- as.integer(table(y))
      u <- (cumsum(u)/sum(u))[1:q]
      zetas <-
         switch(method,
                "logistic"= qlogis(u),
                "probit"=   qnorm(u),
                "cauchit"=  qcauchy(u),
                "cloglog"=  -log(-log(u)) )
      s0 <- c(rep(0,pc),zetas[1],log(diff(zetas)))

##         # try logistic/probit regression on 'middle' cut
##         q1 <- length(lev) %/% 2
##         y1 <- (y > q1)
##         X <- cbind(Intercept = rep(1, n), x)
##         fit <-
##             switch(method,
##                    "logistic"= glm.fit(X, y1, wt, family = binomial(), offset = offset),
##                    "probit" = glm.fit(X, y1, wt, family = binomial("probit"), offset = offset),
##                    ## this is deliberate, a better starting point
##                    "cloglog" = glm.fit(X, y1, wt, family = binomial("probit"), offset = offset),
##                    "cauchit" = glm.fit(X, y1, wt, family = binomial("cauchit"), offset = offset))
##         if(!fit$converged)
##             stop("attempt to find suitable starting values failed")
##         coefs <- fit$coefficients
##         if(any(is.na(coefs))) {
##             warning("design appears to be rank-deficient, so dropping some coefs")
##             keep <- names(coefs)[!is.na(coefs)]
##             coefs <- coefs[keep]
##             x <- x[, keep[-1L], drop = FALSE]
##             pc <- ncol(x)
##           }
##         spacing <- logit((1L:q)/(q+1)) # just a guess
##         if(method != "logistic") spacing <- spacing/1.7
##         gammas <- -coefs[1L] + spacing - spacing[q1]
##         thetas <- c(gammas[1L], log(diff(gammas)))
##         s0 <- c(coefs[-1L], thetas)
    } else if(length(start) != pc + q)
	stop("'start' is not of the correct length")
    else {
        s0 <- if(pc > 0) c(start[seq_len(pc+1)], log(diff(start[-seq_len(pc)])))
        else c(start[1L], log(diff(start)))
      }
    res <- optim(s0, fmin, gmin, method="BFGS", hessian = Hess, ...)
    beta <- res$par[seq_len(pc)]
    theta <- res$par[pc + 1L:q]
    zeta <- cumsum(c(theta[1L],exp(theta[-1L])))
    deviance <- 2 * res$value
    niter <- c(f.evals=res$counts[1L], g.evals=res$counts[2L])
    names(zeta) <- paste(lev[-length(lev)], lev[-1L], sep="|")
    if(pc > 0) {
        names(beta) <- colnames(x)
        eta <- offset + drop(x %*% beta)
    } else eta <- offset + rep(0, n)

    cumpr <- matrix(pfun(matrix(zeta, n, q, byrow=TRUE) - eta), , q)
    fitted <- t(apply(cumpr, 1L, function(x) diff(c(0, x, 1))))
    dimnames(fitted) <- list(row.names(m), lev)
    fit <- list(coefficients = beta, zeta = zeta, deviance = deviance,
                fitted.values = fitted, lev = lev, terms = Terms,
                df.residual = sum(wt) - pc - q, edf = pc + q, n = sum(wt),
                nobs = sum(wt),
                call = match.call(), method = method,
		convergence = res$convergence, niter = niter, lp = eta)
    if(Hess) {
        dn <- c(names(beta), names(zeta))
        H <- res$hessian
        dimnames(H) <- list(dn, dn)
        fit$Hessian <- H
    }
    if(model) fit$model <- m
    fit$na.action <- attr(m, "na.action")
    fit$contrasts <- cons
    fit$xlevels <- .getXlevels(Terms, m)
    class(fit) <- "polr"
    fit
}
```

# Introduction

Recent international crises have experienced a seeming similarity in adversaries responding to their opponent's military actions with dissimilar means. Hamas' cyber attacks in 2019 were met with Israeli airstrikes [@newman_whatisraelstrike_2019], in 2020 Azerbaijani tanks clashed with Armenian drones in the Nagorno-Karabakh region [@dixon_azerbaijandronesowned_2020], Chinese malware attacks have recently joined the fray in an ongoing border skirmish with India [@sanger_chinaappearswarn_2021], and Russian electoral inference in 2016 was met with no concrete response due to concerns about the appropriateness of various kinetic and non-kinetic military operations [@healey_notcyberdeterrence_2018]. There is little consensus about whether to sound the alarm about dangerously escalatory cross-domain interactions [@dixon_azerbaijandronesowned_2020] or cautioning against excess concern [@borghard_israelrespondedhamas_2019]. Furthermore, are these and other recent cases emblematic of the future of warfare [@gatopoulos_nagornokarabakhconflictushering_2020], or is this nothing more than a familiar story, retreaded [@gannon_shadowdeterrencewhy_2021]?

While recent events have motivated an interested in understanding the domains in which states take military action and the consequences those choices have for international stability, this phenomenon has not been studied systematically. Part of this is due to difficulty identifying military domains --- is it simply geographic distinctions between land, air, and sea that match up with different military commands? Where do new tools or technologies like nuclear, space, and cyber fit into this typology? The second difficulty concerns limited empirical data concerning the conduct of conflict. While scholars have developed numerous detailed datasets on conflict's participants, duration, location, and outcome in order to understand war's causes and consequences, empirically-oriented work concerning the conduct of conflict has only more recently received attention [@lindsay_crossdomaindeterrencestrategy_2019]. Evidence is inconsistent since we typically frame it around emerging techs with little empirical evidence [@sechser_emergingtechnologiesstrategic_2019; @horowitz_emergingmilitarytechnologies_2020] or individual domains in isolation [@allen_understandingimpactair_2017; @kostyuk_invisibledigitalfront_2019; @cunningham_maritimerungescalation_2020].

This paper takes an inductive, data-driven approach to identify spatial and temporal patterns in the military domains in which states operate during conflict as well as the relationship between cross-domain interactions and the intensity and duration of international crises.^[On the decision to approaching new social science data inductively, see @grimmer_machinelearningsocial_2021.] Rather than come up with a new typology of military domains, we take a commonly agreed upon understanding shared by practitioners and scholars that contains the traditional domains of land, air, and sea as well as the recent domains of space, cyber, and weapons of mass destruction (WMD).^[While WMD is rarely a distinct military branch and has geographic overlap with the other domains, they do represent a distinct domain in how actors think about them in the international context.] Our motivation in gathering data on the conduct of conflict is driven by the recognition that the domains in which states fight are a reflection of actors' goals, priorities, resolve, and capability and subsequently shape important events like the severity and duration of conflict [@lindsay_politicsmanyother_2020]. In doing so, we produce a novel dataset of the military domains in which `r nrow(df_allnew)` crisis actors operated during `r nrow(df_crisis)` international crises from `r min(df_crisis$yrtrig)` to `r max(df_crisis$yrtrig)` as an expansion to the existing International Crisis Behavior dataset [@brecher_studycrisis_2000].

Our research contains two initial findings that contributing to existing research on the consequences of the domains in which states fight. First, cross-domain conflict is prevalent, but not novel. Some degree of cross-domain military actions occur in `r round(sum(df_crisis$crossdomain_jaccard > 0)/nrow(df_crisis), 2) * 100`% of crises over the past century, with `r round(sum(df_crisis$crossdomain_jaccard == 1)/nrow(df_crisis), 2) * 100`% of crises involving completely dissimilar military actions by belligerents. Moreover that is not a recent phenomenon, as the rate of "cross-domainness" has remained constant over the past century despite the advent of new military domains during that time period like WMD, space, and cyber. Second, contemporary optimism concerning cross-domain conflict may have historical support. Crises in which belligerents engage in cross-domain military conflict are less violent and no more durable than crises in which states respond with like-means. While data on the more salient new domains like cyber remains limited, there is insight to be gleaned about cross-domain interactions in more traditional military domains that should inform theories of deterrence and spiral models of conflict that are being applied to more recent cases.

This paper will proceed in six parts. Section 2 identifies the existing state of the art concerning the domains in which states fight. Section 3 then identifies the theoretical basis underlying both the cross-domain pessimists (cross-domain conflicts are concerning) and optimists (cross-domain conflicts are not concerning) as applied to the intensity and duration of international crises. Section 4 provides an empirical test of these contrasting theories by introducing a novel dataset of the military domains in which states operated during international crises of the past century. Section 5 discusses the implications of these findings for theories of deterrence and the means of conflict, their application to contemporary foreign policy decisions like A2/AD and AirSea battle in East Asia, Russia's "little green men" in Eastern Europe, and kinetic responses to cyber attacks in the Middle East, and avenues for future research using new data on the military domains of international crises. Section 6 concludes.

# Existing theories of conflict escalation and how states fight
The vast array of theorized determinants of conflict escalation -- defined here in terms of its intensity and scope -- primarily concern factors present before war occurs [@morgan_untyingknotwar_1994; @fearon_rationalistexplanationswar_1995; @powell_stabilitydistributionpower_1996]. Since most wars end with an agreement rather than a 'fight to the death', the conduct of conflict, whether actual or latent, has much to tell us about when that war will escalate [@gartner_openingblackbox_1998; @wagner_bargainingwar_2000; @biddle_militarypowerexplaining_2005; @boot_warmadenew_2006; @biddle_strategywar_2007] and how long it will last [@slantchev_howinitiatorsend_2004].

There is a well developed literature on how different \textit{ways} conflicts are fought might influence conflict escalation.^[The conduct and consequences of conflict has also been linked to one another since various states characteristics influence the militaries they choose to develop [@kennedy_risefallbritish_1976 177-204].] The ways in which militaries fight influences war's participants [@fordham_verysharpsword_2004; @beckley_emergingmilitarybalance_2017], victors [@rosen_winningnextwar_1991; @lyall_ragemachinesexplaining_2009; @cappellazielinski_organizingperformancecoalition_2020], costs [@caverley_democraticmilitarismvoting_2014], severity [@talmadge_emergingtechnologyintrawar_2019; @gartzke_influenceseapower_2020], location [@schilde_europeanmilitarycapabilities_2017; @crisher_navalpowerendogeneity_2017], and duration [@martinezmachain_aircampaignduration_2015; @cappellazielinski_politicaleconomynational_2016; @caverley_militarytechnologyduration_2017] as well as what states are able to project power [@corbett_principlesmaritimestrategy_1911; @beasley_closingpresencegap_2015], what issues are resolved with force [@allison_armamentsarmscontrol_1975; @becker_clearingairtransatlantic_2017; @fearon_cooperationconflictcosts_2018; @sechser_emergingtechnologiesstrategic_2019], what threats are credible [@buzan_armsdynamicworld_1998; @slantchev_powerhurtcostly_2003; @post_flyingfailcostly_2019; @montgomery_signalsstrengthcapability_2020], the efficacy of states' reputation [@erickson_dangeroustradearms_2018], and the balance of power [@glaser_politicalconsequencesmilitary_1992; @horowitz_diffusionmilitarypower_2010; @gartzke_determinantsnuclearforce_2014].

Despite interest in the means by which state engage in conflict, there is little agreement conceptualizing and categorizing those means. One broad way to conceptualize them is in terms of the domains in which they take place. There are many ways of thinking about domains, as they can be distinguished from one another by technology, tactic, geography, or purpose [@lindsay_crossdomaindeterrencestrategy_2019]. Some have questioned the utility of conceptualizing these domains as distinct, especially as new domains like space and cyber increasingly play the role of supplementing rather than supplanting operations in the more traditional domains [@stallone_donforgetcyber_2009; @caton_domainscommonscontext_2012; @libicki_cyberspacenotwarfighting_2012]. Rather than engage in ontological debates about the best way to typologize state military behavior, we take domains here as a useful starting point for thinking about differences in the way states fight. We know that practitioners think about these domains as distinct and we have theories about the advantages and disadvantages of taking military action in each of these domains [@lindsay_crossdomaindeterrencestrategy_2019]. I use the term domain here loosely to refer to land, air, sea, WMD, space, and cyber.^[Zooming out, international actors, of course, interact in non-military ways. Some research thinks about this more broadly, where military interactions represent a domain among other strategies that are economic and diplomatic. I bracket that discussion here since the goal of this paper is to contribute to research that disaggregates that military domain. For existing work on issue linkages involving military and non-military domain interactions, see @rosenau_comparativeforeignpolicy_1968, @tollison_economictheorymutually_1979, @vasquez_tangibilityissuesglobal_1983, @morgan_issuelinkagesinternational_1990, @morgan_modelforeignpolicy_2000, @starr_substitutabilityforeignpolicy_2000, @poast_doesissuelinkage_2012, @greenhill_asymmetricadvantageweaponizing_2019, and @huang_militarydeterrencemultidimensionality_2019.] I choose these domains because they differ from one another in how political actors think about them and because they have different constraints concerning concepts of interest like power project capabilities, movement, coordination, casualty risk, and cost [@lindsay_politicsmanyother_2020, 9-10].^[Zooming in, there are differences in the strengths and weaknesses of capabilities that produce unique cross-domain dynamics regarding conflict intensity and duration [@lindsay_politicsmanyother_2020]. We bracket that discussion here given limitations in empirical data, but it represents a fruitful avenue for future research.]

If the problem of the security dilemma is to decide whether a particular response to an adversaries behavior will deter or escalate, the problem as applied to the conduct of conflict is whether the military means one chooses determine the likelihood of escalation. @snyder_balancepowerbalance_1965 [187] described the importance of new (nuclear) means and escalation long ago. “[N]uclear technology introduced a new form of intent-perception and a new form of uncertainty — that concerning what types of military capability the opponent was likely to use and what degree of violence he was willing to risk or accept”. This view is widely credited with originating discussion of the stability-instability paradox regarding the relationship between capabilities an actor possesses and the capabilities they use. The resulting debate created competing perspectives on escalation, characterized as deterrence and the spiral model. Snyder’s deterrence perspective contrasts with the spiral model advocated by @jervis_illogicamericannuclear_1984. @powell_nuclearbrinkmanshiplimited_2015's recent work represents the newest attempt to theorize the relationship between power and risk. He develops a formal model where the challenger decides how much military power to use to achieve its ends. The more power it uses, the higher its chance of winning, but the higher the potential risk of escalation as well. Although an important contribution to understanding conflict escalation and the deterrence and spiral models of conflict, what it means to bring "more power to bear" in a crisis is unspecified [@powell_nuclearbrinkmanshiplimited_2015, 598].

Just like the amount of capabilities a state devotes to a conflict influence its outcome, the nature of those capabilities similarly matters [@carcelli_diversificationdeterrencenew_2017]. In one sense, it is intuitive to note that military domains have some inherent level of belligerence that can be compared to one another. A nuclear strike is likely to be interpreted as more intense belligerence than a non-nuclear attack at any feasible level of use. Yet this intuition remains coarsely formulated. Foundational conceptions of escalation ladders @kahn_thermonuclearwar_2007 have evolved in a way that more closely parallels existing "domains" like @libicki_cyberdeterrencecyberwar_2009 [28-29] noting belligerence increases as an actor shifts from diplomatic and economic to cyber, then physical force, then nuclear force. But this is coarse, since not all conventional domains involving physical force are created equal and more importantly it does not account for the strategies in which the other side is operating. Whether the move from one "physical force" domain to another is consistent with the deterrence or spiral model of conflict remains unknown.

With this in mind, evaluations of the importance of actions in domains has recently shifted to theorizing about interactions across domains [@lindsay_introductioncrossdomaindeterrence_2019]. There is suspicion that cross-domain interactions are more prevalent now because there are new domains in which states can operate. While literature on the nuclear domains and conventional-nuclear interactions is no longer in its infancy [@kissinger_limitedwarconventional_1960; @wallace_armsracesescalation_1979; @weede_armsracesescalation_1980; @carver_conventionalwarfarenuclear_1986], this has recently expanded to cross-domain or multi-domain interactions involving newer domains of space [@early_exploringfinalfrontier_2014; @lin-greenberg_privateeyessky_2021], cyber [@healey_notcyberdeterrence_2018; @libicki_cyberspacenotwarfighting_2012; @nye_deterrencedissuasioncyberspace_2017; @schneider_informationrevolutioninternational_2017; @lindsay_coercioncyberspacestabilityinstability_2018; @kostyuk_invisibledigitalfront_2019], drones [@boyle_costsconsequencesdrone_2013; @byman_whydroneswork_2013; @hazelton_dronestrikesgrand_2016; @horowitz_consequencesdroneproliferation_2016; @schneider_whytroopsdon_2017; @zegart_cheapfightscredible_2018], artificial intelligence (AI) [@jensen_algorithmswarpromise_2020; @johnson_delegatingstrategicdecisionmaking_2020], and quantum computing [@lindsay_demystifyingquantumthreat_2020; @gamberini_quantumsensingpotential_2021]. Research has expressed concern that these and other new tools of conflict may allow new forms of insidious revisionism [@brands_paradoxesgrayzone_2016], although skepticism about the novelty and efficacy of these "new" forms of warfare casts doubt on those concerns [@gannon_shadowdeterrencewhy_2021].

Although these new domains have motivated an increasing interest in the subject, empirical evidence remains limited since many of the domains of interest (e.g. cyber) have less well established evidence of empirical trends [@horowitz_emergingmilitarytechnologies_2020]. Even among recent attempts to empirically identify the ownership and use of new modes of warfare like satellites [@early_exploringfinalfrontier_2014 on satellites], cyber attacks [@valeriano_dynamicscyberconflict_2014; @kostyuk_invisibledigitalfront_2019], drones [@horowitz_consequencesdroneproliferation_2016], and smart bombs [@kahn_whogetssmart_2021], the efforts are contained to a single capability or domain. As a result, existing research on cross-domain conflict is it tends to think about the "cross" component being across domains within an actor rather than across domains across actors. @nye_deterrencedissuasioncyberspace_2017 [46] recognizes this in noting "it is a mistake to see the cyber realm in isolation. The term “cyber deterrence” can be confusing because theorists tend to focus on in-kind or in-domain deterrence rather than on a broad range of tools that can be used both actively and passively and with graduated effects. A response to a cyberattack need not be by cyber means any more than a response to a land attack need be by the army rather than naval or air forces."

We agree with @powell_nuclearbrinkmanshiplimited_2015 [598] when he states “the amount of power the challenger brings to bear affects the stability of the conflict. More specifically, how much power the challenger brings to bear limits how much risk the defender can generate”. We expand upon his conceptualization of the military might used by the challenger. Powell varies "military might" as different types or levels of conflict [@powell_nuclearbrinkmanshiplimited_2015, 599]. But types and levels of conflict are not synonymous, nor should they be grouped together [@kreps_escalationfirebreakscyber_2019]. We seek to provide insight about how the relationship between the types of conflict by adversaries explains differences in conflict escalation.

# Is a ship a fool to fight a fort?
Royal Navy Admiral Nelson is (mis)credited with having proclaimed "a ship's a fool to fight a fort" [@ferreiro_horationelsonnever_2016]. Interpreting foolish actions to be those that escalate the conflict in violence and/or duration, we can derive two contrasting expectations about cross-domain interactions and conflict escalation. When the conduct of conflict involves dissimilar military domains --- as in the recent cases of Azerbaijan-Armenia, China-India, Israel-Hamas, and US-Russia --- will that cross-domain interaction deter escalation, or cause it to spiral? Despite opposing predictions, both theories share the assumption that the manner in which states engage in a crisis influences whether it is resolved peacefully or violently since it influences how states show resolve and/or capability. These theories thus differ in whether demonstrations of resolve and capability succeed or fail when they involve taking actions in a military domain in which your opponent is not operating.

## Cross-domain escalation: the danger of an ear for an eye
Crises in which actors interact in dissimilar military domains may be more violent and longer because cross-domain interactions complicate interpretations of proportionality and stake, thus contributing to misinterpretation of resolve [@morrow_internationallawcommon_2019]. If a belligerent taking action with 100 ground troops is met with the defender deploying 200 ground troops, the belligerent could reasonable interpret the defenders action as a "re-raise" indicating the defender places high stakes on the issue in dispute. But if the defender responds to a deployment of 100 troops with 5 aircraft, it is less clear whether that is a re-raise. In essence, cross-domain conflict presents actors with an apples to oranges comparison concerning the resolve and value for the issue at stake that makes a negotiated settlement more difficult to come by given ambiguity about the extent of the bargaining range.

Uncertainty can cause war by creating incompatible expectations about war's utility and outcome [@fearon_rationalistexplanationswar_1995; @gartzke_warerrorterm_1999]. Asymmetry in the military domains in which belligerents are taking actions can be a source that uncertainty. Cross-domain interactions may be perceived by the adversary as shifting to a domain where one has escalation dominance because you highly value the stakes of the conflict. Since each military domains has different strengths and weaknesses concerning outcomes like credibility, warfighting, costs, and signaling, operating in a different domain that your adversary may signal a difference in one's tolerance for cost and risk [@carcelli_diversificationdeterrencenew_2017; @gartzke_signalingforeignpolicy_2017; @lindsay_crossdomaindeterrencestrategy_2019]. A stylized representation of military domains is as a game of rock, paper, scissors where each domain's characteristics bring advantages against some of your opponent's chosen domains, but disadvantages against others [@lindsay_crossdomaindeterrencestrategy_2019 16, 350]. In discussing the dangers of Chinese power projection from land, former US National Security Adviser H.R. @mcmaster_harbingersfuturewar_2016 advised practitioners "recognize that synergy, right, between joint force capabilities and how, you know, really joint operations - it's rock, paper, scissors, you know? So if you can - if you have all of those tools available - maritime, aerospace, cyberspace, land capabilities - then you're able to pose that enemy with multiple dilemmas." While responding to your adversaries rock with paper may help you win the conflict, it could also produce an incentive for your opponent to augment their hand with scissors, thus encouraging escalation. Responding in-kind --- playing rock against rock --- may be de-escalatory because it represents a symbolic gesture to opt against an efficient response (paper) that would require the opposing side to engage in a new, more efficient response (scissors).^[A similar argument about the defense-dominant nature of responding in-kind is made by @biddle_futurewarfarewestern_2016.] Responding in-kind thus represents a symbolic gesture to opt against an efficient response that would require the opposing side to engage in a new more efficient response. By responding in-kind, states can agree to call it a draw in a way that represents a face saving measure [@carson_facingsavingface_2016; @carnegie_spotlightharshglare_2018; @carson_secretwarscovert_2018]. Conversely, undertaking that engagement in a new military domain may unintentionally create an "escalatory updraft" if the opponent misinterprets what should be a tit-for-tat response that ends aggression with one that instead escalates it [@oneill_conflictualmovesbargaining_1991, 104]. To return to @powell_nuclearbrinkmanshiplimited_2015's logic about bringing more power to bear, taking action in a new domain may represent a reduction in cost for you, but the very logic of escalation dominance that generates that outcome also means interacting in the new domain represents an increase in cost for your opponent, with adverse consequences for the likelihood of a peaceful settlement.

Scholars have raised this concern in the context of the US and China in arguing "A war at sea could thus quickly become a war on land, potentially even raising risks of nuclear escalation if the US starts to erode potential capabilities relevant to China's nuclear arsenal" [@talmadge_emergingtechnologyintrawar_2019, 880-881].^[See also @talmadge_wouldchinago_2017 and @cunningham_maritimerungescalation_2020 for more on China and the risk of nuclear escalation.] States often share an assumption that opponents will "follow precedent by responding in kind with similar weapons against a similar target set", so failing to do so risks un-calibrated escalation [@warden_limitednuclearwar_2018, 24]. In the cyber context, scholars have argued retaliating to physical force with a cyber attack may raise issues of proportionality [@libicki_cyberdeterrencecyberwar_2009]. While China thinks that cyber attacks are a proportional response to US trade pressure given the stakes each place on the sectors being harmed by the other actions, the United States disagrees. "The U.S. approach to Chinese gray zone tactics seems to be consistently several steps behind the threat. As in the Russia case, this is likely due in large part to the absence of a coherent strategic approach, leading to ad hoc U.S. responses from one crisis point or domain of interaction to the next." [@hicks_othermeanspart_2019, 8].

## Cross-domain deterrence: the safety of playing chess while your opponent is playing checkers
The opposite logic could also be at play. Crises involving states taking military actions in dissimilar domains could be less likely to escalate by communicating a change in the stakes and a willingness to escalate or by giving one side the upper hand if conflict actually breaks out [@rovner_giveinstabilitychance_2020]. The apples to oranges analogy may be backwards; rather than creating confusion about resolve and the opponent's evaluation of the stakes, a willingness to play a different game entirely may sharpen that logic in a way that signals it may be better to try to resolve the conflict without fighting.

Rather than simply thinking of military domains as playing rock, paper, or scissors, the virtues of different domains in their ability to signal or fight wars also creates distinctions in what it means to "play" one of them. Threats differ from forward deployments which are further different from seizing territory or discharging a rocket [@lai_effectsdifferenttypes_2004; @mackey_prewarmobilizationtermination_2014; @post_flyingfailcostly_2019]. In this way, operating in a new domain could deter rather than inflame by providing an avenue to graduated escalation [@cashman_introductioncauseswar_2007; @montgomery_signalsstrengthcapability_2020]. Reinforcing an army unit by putting bombers on alert or forward deploying a naval unit could be a graduated way of signaling resolve and a willingness to escalate without necessarily increasing the number of army troops already on the battlefield [@slantchev_militarycoercioninterstate_2005; @slantchev_militarythreatscosts_2011; @warden_conventionalnuclearintegrationnext_2020]. Using the same assumption of escalation dominance, action in a new domain in which your opponent is not operating because they do not have escalation dominance in that domain can be a costly signal that makes a negotiated settlement more likely [@cimbala_militarypersuasiondeterrence_1994; @quek_arecostlysignals_2013; @handberg_aircraftcarrierballistic_2016]. In the same way that children on a playground may yell 'hotter' or 'colder' as their blindfolded classmate tries to find some object, operating in new domains can add credibility to verbal statements about whether an actor perceived an offered negotiation as preferable to continuing to fight. Furthermore, deterrence may work not because of mutual vulnerability, but because of one-sided fear about overwhelming punishment [@green_revolutionthatfailed_2020; @lieber_mythnuclearrevolution_2020]. If a state operating in a new domain during a crisis is indicative of willingness to take action in a domain where they have escalation dominance, that may convey to the other side that the state is willing to incur a higher cost.

Historically, Nixon placing bombers on nuclear alert during the 1969 Giant Lance operation demonstrated the introduction of a new domain as a way to signal a willingness to gradually escalate or de-escalate based on their opponent's actions [@sagan_madmannuclearalert_2003; @burr_nixonadministrationhorror_2005]. This case is consistent with existing theories of signaling in finding that some domains are better suited to turning the dial incrementally rather than all at once [@kroenig_logicamericannuclear_2018; @post_flyingfailcostly_2019; @montgomery_signalsstrengthcapability_2020]. President Kennedy had a similar rationale behind the "flexible response" doctrine in 1961, which aimed to provide more options for controlled escalation. In the event of more limited Soviet aggression like an attempt to hold Berlin hostage, proponents of flexible response hoped the more credible threat of responding with other conventional forces would be sufficient to get the Soviet Union to back down [@duffield_evolutionnatostrategy_1991].^[There is disagreement concerning whether Kennedy actually believed in --- and would have implemented --- "flexible response" in the ways espoused at the time or if it merely served political purposes [@gavin_mythflexibleresponse_2001]. We use this example here to illustrate the strategic logic held by proponents of the policy, rather than if the United States would have actually followed through.] Similarly, a ceasefire was declared between Israel and Hamas the day after Israel responded to Hamas' cyberattacks with an airstrike [@morris_israelgazamilitants_2019]. One way to interpret this event is the very conditions that make a cross-domain response an immediate escalation make it a medium and long-term de-escalation. By showing that an opponent has gone over the line in the intensity of their aggression, an unexpected cross-domain response could convey that there are significant costs to be had if they do not pull back.

As an aside, cross-domain interactions may not be the direct cause of escalation (or de-escalation), but rather an observable indicator for different motives to escalate in a way that masks the stakes [@slantchev_militarythreatscosts_2011]. States with more at stake may be more willing to escalate, but cross-domain interactions may communicate that poorly since it relies on your opponent identifying the costs of your action when that same action would generate different costs for them.^[For existing research on motives for escalation, see @lebow_whynationsfight_2010, @dafoe_reputationstatusmotives_2014, @paul_statusworldpolitics_2014, @fikenscher_seekingstatusrecognition_2015, @renshon_statusdeficitswar_2016, @lupton_signalingresolveleaders_2018, and @shifrinson_linkagepoliticsmanaging_2019.] Rather than try to empirically identify whether the underlying causes of cross-domain interactions are directly or indirectly determinants of conflict intensity, we simply hope to shed light on the association between the use of these domains --- whatever their origins --- and crisis outcomes.

# Empirics
## Research Design
This paper's primary empirical contribution is the creation of a new dataset recording the domains in which warfighting and military threats occurred during `r nrow(df_crisis)` distinct crises from `r min(df_crisis$yrtrig)` to `r max(df_crisis$yrtrig)`. In sum, the new data contains information on `r nrow(df_allnew)` crisis actors, the majority of which are states.^[International organizations and non-state actors are often involved in crises and sometimes deploy military or quasi-military capabilities like peacekeepers or foreign aid distributors. We do not consider the military logic of these actors here, but the data is available for future work on the topic.] This crisis-domain dataset represents -- to the best of our knowledge -- the most extensive data collection available on the domains in which militaries operate during conflict.

Doing so involved three distinct steps. First, we event coded the `r nrow(df_crisis)` crises in the International Crisis Behavior (ICB) dataset [@brecher_studycrisis_2000; @brecher_internationalcrisisbehavior_2020], identifying the military domains in which crisis actors acted at the sentence level. Second, we re-organized and subset that data to the crisis-dyad level using pre-existing work on ICB crisis dyads. Third, we create a new measure of the dissimilarity of the domains in which each side in the crisis took military actions, which we will refer to as "cross-domainness".

### Event coding
We extend the ICB dataset by gathering extensive data on the tools utilized during crises. The ICB data provides a unique opportunity to examine conflict escalation because it is not limited to cases where conventional military conflict occurred. Rather, the data examines crises -- some of which escalated to militarized interstate conflict with some degree of violence and others which did not. As a result, this data provides variation in the dependent variable since we can compare cases where a crisis did violently escalate to cases where the crisis did not. Examining patterns in the military domains and units involved in each of these cases can shed light on when a crisis is likely to escalate. The dataset introduced here contains data on the military domains used by each actor during these international crises. As such, the unit of analysis is the crisis-actor and the newly coded variables are binary true/false values representing whether that actor employed that domain during the crisis.

The data source used by the coders was the ICB crisis narratives that provide qualitative descriptions of the crises. This has a few benefits over other events datasets. First, the ICB crisis narratives are written in a systematic fashion. They are all approximately the same length, contain the same level of detail, and were written by the same research team at the University of Maryland. This reduces concerns that variation among crisis variables is due to variation in the data generating process. For example, more recent crises do not have more detailed crisis narratives which is not the case for the reporting of militarized interstate disputes captured by scraping news sources. Second, the ICB narratives are accompanied by the more commonly used ICB dataset that codes important international relations variables in the dataset. This helps verify the coding effort since the information that coders extract from the crisis narratives should be consistent with previous codings that populate the quantitative ICB dataset.

The data created here involves the domains in which states took actions during international crises.^[This is a subset of a broader project producing event data using the ICB narratives. For the complete event data and more detailed explanation of the underlying ontology, see @carcelli_introducinginternationalcrisis_2021.] We define the domains in which a crisis-actor can take a military action based on the military units that undertook the action, as described in Table \@ref(tab:domains). The coders distinguished actions, speech acts, and thoughts for consistency with pre-existing event datasets like CAMEO and Phoenix [@schrodt_cameoconflictmediation_2005; @schrodt_precedentsprogressprospects_2012; @althaus_clinecenterhistorical_2020]. We define actions as a physical act performed by one of more actors. Examples of actions include raises in alert level, mobilizations, fortification, military exercises, weapons tests, deployments, shows of force, blockades, border violations, attacks, invasions, and bombardments. Planning to take an action does not constitute an action unless this action is subsequently carried out. For example, a state making a verbal threat to send tanks into a neighboring country is not is not coded as a ground action unless the state mobilized, deployed, invaded, attacked or otherwise took a subsequent physical action with those tanks. 

\begin{table}[]
\centering
\begin{tabular}{|l|l|}
\hline
\textbf{Domain} & \textbf{Description} \\
\hline
Air & Bombers, fighters, and missiles \\
Cyber & Information operations and cyber disruption \\
Land & Armored vehicles, artillery, and troops \\
Sea & Aircraft carriers, submarines, and surface ships \\
Space & Satellites and surveillance beyond the earth's atmosphere \\
WMD & Nuclear, chemical, and biological weapons \\ 
\hline
\end{tabular}
\caption{Military domains coded for each crisis actor. Codings are binary with 1 indicating the actor took an action in that domain during the crisis and 0 otherwise.}
\label{tab:domains}
\end{table}

The first set of codings were done by graduate students at the University of Maryland that were overseen by the ICB directors who helped construct the original ICB datasets and narratives. For each crisis, two research assistants coded the ICB crisis narratives at the sentence level for a new series of variables. Coders could not make reference to the existing codings and the raw coding includes a measure of uncertainty for each coder. Coders selected one of the `r nrow(df_crisis)` ICB crises and for each individual sentence in the corresponding ICB crisis narrative, they coded the events that occurred as a series of actions undertaken by an actor along with supporting details like when the action took place, where, and to what effect. The codings were then double checked, with particular attention paid to the less common domains of WMD, space, and cyber. The final version of the data includes an aggregation of the multiple different codings that exist for each case. This paper only focuses on the military domains used during these crises. This represents an improvement over current efforts to analyze aggregate analysis of crisis behavior by including fine-grained information about how actors behaved. Figure \@ref(fig:domain-count) shows the distribution of actions taken in each domain by crisis actor. Consistent with expectations, the past century witnessed crisis-actors most often operating in the land domain, with significantly fewer WMD, cyber, and space actions taking place. Even so, the prevalence of WMD events may appear higher than expected. This is because the domain variables refer to where the action took place. This does not refer to what was used during the action, but where the action happened. As a result actions like raising nuclear alert levels or forward deploying nuclear bombers constitute WMD "actions" even if nuclear bombs were not subsequently detonated. The low number of space and cyber actions do not reflect the rarity of military action in the space and cyber domains, but rather the rarity of those domains \textit{in international crises}. This distinction is important, as existing work has documented numerous cases of cyber attacks that occur outside the universe of international crises defined by the International Crisis Behavior project [@valeriano_dynamicscyberconflict_2014].

```{r domain-count, fig.cap = "Distribution of domains by crisis-actor", fig.width = 4, fig.height = 4}
df_allnew %>% 
  dplyr::select(dplyr::ends_with("_action")) %>%
  dplyr::rename_at(vars(dplyr::ends_with("_action")),
                   dplyr::funs(stringr::str_replace(., "_action", ""))) %>%
  tidyr::gather() %>%
  dplyr::mutate(key = stringr::str_to_title(key)) %>%
  dplyr::group_by(key) %>%
  dplyr::filter(!is.na(value)) %>%
  dplyr::summarise(value = sum(value == 1)) %>%
  ggplot(aes(x = reorder(key, -value), 
             y = value)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = value, 
                y = value + 20), 
            vjust = 0, 
            size = 5
            ) +
  ylim(0, 1500) +
  labs(title = "Crisis-Actors per Domain (1918-2015)",
       x = "Domain",
       y = "Number of Crisis-Actors") +
  theme_bw() +
  theme(panel.grid = element_blank(), 
        text = element_text(size = 12)
        )
```

Of course, actors do not always restrain their military activities to a single domain. Conventional wisdom surrounding the efficacy of full-spectrum military forces lead one to rightly believe that actors often engage in multiple military domains simultaneously, especially as the stake of a conflict heighten. Figure \@ref(fig:domain-combos) describes the combination of military domains that each crisis-actor undertook. While the sole deployment of land forces still remains by far the most common form of military action, combined land- and air operations is the second most common, followed by land, sea, and air being used in unison.

```{r domain-combos, fig.cap = "Combination of domains for each crisis actor. Each bar represents the number of crises with that unique intersection of domain values present", fig.width = 5, fig.height = 3.5}
df_actor %>% 
  dplyr::select(dplyr::ends_with("_action")) %>%
  dplyr::rename_at(vars(dplyr::ends_with("_action")),
                   dplyr::funs(stringr::str_replace(., "_action", ""))) %>%
  janitor::clean_names("title") %>%
  UpSetR::upset(number.angles = 0, 
                point.size = 2.5, 
                line.size = 1,
                mainbar.y.label = "Number of Crisis-Actors", 
                order.by = "freq", 
                set_size.show = TRUE, 
                set_size.scale_max = 1250, 
                sets.x.label = "Domain Count", 
                text.scale = c (1.5, 1.5, 1.5, 1, 1.5, 1.5))

```

### Identifying crisis-dyads
The original ICB data exists at two levels of analysis, the crisis-level and the crisis-actor level. A crisis is defined as an international event that meets 3 conditions. First, an actor must perceive a threat to one or more of its basic values. Second, the actor must have an awareness of a finite timeline for responding to that value threat. And third, there must be a heightened probability of military hostilities. A crisis can escalate to an actual military dispute, but that does not always happen. Many crises are resolved prior to escalation to actual conflict because the existence of the crisis motivates diplomatic solutions to resolve the crisis without resorting to war. While the actor-level data latter contains more granularity, it leaves the role of each actor unspecified. So after identifying the military domains in which crisis-actors took actions, this new data was merged with existing data on ICB crisis dyads [@hewitt_dyadicprocessesinternational_2003; @beardsley_winningbomb_2009; @levin-banchik_precrisismilitaryhostility_2020]. Three conditions determine whether an actor is included is the ICB crisis-dyad data. First, both sides of the dyad must be states as defined by the @gleditsch_revisedlistindependent_1999 state list dataset. Second, at least one state must meet all the crisis conditions specified by the original ICB project. Third, at least one actor must perceive that the other has directed a threat or hostile action against them.

Although the new ICB domains data created for this project contains information on `r nrow(df_allnew)` crisis-actors, not all of them are members of crisis-dyads. After subsetting to the crisis-dyad conditions, we are left with `r nrow(df_actor)` crisis-dyad actors that were on one of two sides in `r nrow(df_crisis)` previously identified crisis-dyads. The loss of `r nrow(df_allnew) - nrow(df_actor)` actors is primarily attributed to newly gathered information on non-state actors (including international organizations like the Organization of American States and United Nations) or actors that were mentioned in the crisis narratives as a mediator or neighboring state, but not a crisis participant involved as one of the two sides in the crisis. Numerous crises involve multiple dyads if, for example, military coalitions were involved. To simplify coding, when multiple actors participated on the same side in the conflict, that side was coded as having taken actions in a given military domains if any actor on that side took actions in that domain. For example, if both France and the United States were coded as being on the same side (side A) in a crisis and France deployed naval assets and the United States deployed ground forces, side A is simply coded as having taken action in both the naval and land domain.

### Measuring cross-domainness
The third and final step involves identifying the dissimilarity of the domains in which each side of each crisis-dyad took actions. We call this measure cross-domainness which is a variable at the crisis level of analysis. For each crisis, if the two sides took actions in identical domains, cross-domainess is low. If the two sides took actions in entirely distinct domains, cross-domainness is high.

We produce this measuring by creating a Jaccard similarity coefficient comparing the military domains in which each of the two sides took actions during each ICB crisis. This measure identifies the union of domains in which each side took actions as a ratio of those in which only one side took action such that for a crisis-dyad with two sides A and B, $J_{(A, B)} = \frac{A \cap B}{A + B - (A \cap B)}$. Table \@ref(tab:jaccard) provides descriptive examples of how domain similarity in each crisis would be measured. During North Korea's nuclear test in 2009, North Korea launched a satellite followed by an underground nuclear test. The United States, Japan, and South Korea moved naval forces as a part of military drills and kept interceptor-capable ships in the region following those drills. Since North Korea took action in the space and WMD domain while the US, South Korea, and Japan responded in the naval domain, this crisis had entirely cross-domain interactions -- the two sides took no action in the same domain. By contrast, the Kashmir crisis that involved India, Pakistan, and China in 1965 only involved land forces being deployed by both sides. India sent troops across the ceasefire line in response to infiltration by Pakistani "freedom fighters" and China then responded with troop movemens near the border. In this case, the two sides responded exclusively "in-kind", so this crisis was entirely within-domain. The Jaccard measure is appropriate because we each crisis contains precisely two sides, meaning there are two vectors to compare, the values are binary, and we are concerned with the similarity of measures that were employed as opposed to those that were not. Similar measures like the simple matching coefficient (SMC) are less appropriate because they weigh mutual 0's as an increase in similarity.^[For an overview comparing different similarity measures and criteria for their appropriate application, see @ahlgren_requirementscocitationsimilarity_2003 and @egghe_goodpropertiessimilarity_2010.] These kinds of measures are less appropriate for measuring similarity in military domains since there is variation across space and time about what 0's mean. In some cases it means omission by choice but in others it means omission by necessity. Although neither the US nor Vietnam used WMDs during the Vietnam war, that doesn't make their military strategy similar because the US could have done so but Vietnam could not have, so those 0's mean different things. We avoid this problem by measuring similarity in terms of what they did use.

```{r}

```

  \begin{table}[]
  \centering
  \begin{tabular}{|l|l|cccccc|c|}
  \hline
    \textbf{Crisis} & \textbf{Belligerents} & \multicolumn{1}{l}{\textbf{Land}} & \multicolumn{1}{l}{\textbf{Air}} & \multicolumn{1}{l}{\textbf{Sea}} & \multicolumn{1}{l}{\textbf{WMD}} & \multicolumn{1}{l}{\textbf{Space}} & \multicolumn{1}{l|}{\textbf{Cyber}} & \multicolumn{1}{l|}{\textbf{Jaccard}} \\ 
    \hline
    \multirow{2}{*}{N Korea Nuclear IV (2009)} & US, Japan, S Korea & - & - & 1 & - & - & - & \multirow{2}{*}{1} \\
     & N Korea & - & - & - & 1 & 1 & - & \\
    \hline
    \multirow{2}{*}{Yemen War IV (1967)} & Yemen, Egypt & 1 & 1 & - & 1 & - & - & \multirow{2}{*}{0.67} \\
      & Saudi Arabia & 1 & - & - & - & - & - & \\ 
    \hline
    \multirow{2}{*}{Kashmir I (1947)} & India & 1 & 1 & - & - & - & - & \multirow{2}{*}{0.5} \\
      & Pakistan & 1 & - & - & - & - & - & \\ 
    \hline
    \multirow{2}{*}{Gulf of Tonkin (1964)} & S. Vietnam, U.S. & 1 & 1 & 1 & - & - & - & \multirow{2}{*}{0.33} \\
      & N. Vietnam & 1 & - & 1 & - & - & - & \\ 
    \hline
    \multirow{2}{*}{Kashmir II (1965)} & India & 1 & - & - & - & - & - & \multirow{2}{*}{0} \\
     & Pakistan, China & 1 & - & - & - & - & - & \\ 
    \hline
  \end{tabular}
  \caption{Example of cross-domainness measure at the crisis-level using the Jaccard similarity coefficient. The measure is bounded between 0 and 1, with 0 indicating the two sides took actions in identical domains and 1 meaning complete dissimilarity.}
  \label{tab:jaccard}
  \end{table}

Figure \@ref(fig:cddness) shows the distribution of cross-domainness in every ICB crisis. This figures demonstrates that cross-domain military crises is not novel --- in fact, it represents the modal form of conflict over the past century. Of the `r nrow(df_crisis)` ICB crises, `r round(sum(df_crisis$crossdomain_jaccard == 1)/nrow(df_crisis), 2) * 100`% are entirely cross-domain, meaning the two side had zero overlap in what military domains they operated in during the crisis, and `r round(sum(df_crisis$crossdomain_jaccard > 0)/nrow(df_crisis), 2) * 100`% had at least some amount of cross-domain interaction. In only `r round(sum(df_crisis$crossdomain_jaccard == 0)/nrow(df_crisis), 2) * 100`% did both sides behave in-kind. Figure \@ref(fig:cddness-time) shows that the temporal distribution of cross-domain interactions runs contrary to conventional wisdom. Despite the common demarcation of the nuclear age and advent of new domains like cyber and space, international crises are not more cross-domain than they used to be. Although new military tools have become available to states, they are either not being utilized in international crises, or they are being utilized only in cases where the opposing side responds in kind.

```{r cddness, fig.cap = "Distribution of cross-domainness in international crises. Higher values represent higher cross-domain interactions between adversaries", results = 'asis', out.width = '80%'}
df_crisis %>%
  ggplot(aes(x = crossdomain_jaccard)) +
  geom_histogram() +
  labs(title = "Cross-domainness Across Crises (1918-2010)",
       x = "Cross-domainness",
       y = "Number of crises"
       ) +
  theme_minimal() +
  theme(text = element_text(size = 14))
```

```{r cddness-time, fig.cap = "Distribution of Cross-domainness over time. Line represents a bivariate generalized linear model with the shaded area corresponding to the 95\\% confidence interval", results = 'asis', out.width = '80%'}
df_crisis %>%
  ggplot(aes(x = yrtrig,
             y = crossdomain_jaccard)) +
  geom_point() +
  stat_smooth(method = "glm") +
  ylim(0, 1) +
  labs(title = "Cross-domainness Over Time (1918-2010)",
       x = "Year",
       y = "Cross-domainness") +
  theme_minimal() +
  theme(text = element_text(size = 14))
```

## Model
The unit of analysis is the international crisis which has been collapsed from the crisis-dyad level using the method described. To test hypotheses about the relationship between cross-domain military interactions and the crisis escalation, we model the relationship using two separate independent variable --- intensity of violence and crisis duration. We choose these two dependent variables because they are common observable indicators for conflict escalation [@allen_timebombsestimating_2007; @asal_proliferationinternationalcrisis_2007]. 

Two different model classes are required to estimate the impact of cross-domainness on intensity of violence and crisis duration. The dependent variable for the first model --- intensity of violence --- is measured on an ordinal 1-4 scale with 1 describing no violence and 4 describing-full scale war. The dependent variable for the second model is crisis duration measured in days. Figure \@ref(fig:dv) shows the distribution of the dependent variables, both of which come from the original ICB crisis-level data. The median crisis duration is `r median(df_crisis$brexit)` days. Because Model 1 has an ordinal dependent variable, the appropriate model specification is an ordered probit [@johnston_generalizedorderedprobit_2020]. Because Model 2 is a duration variable operationalized as the months from crisis initiation to crisis termination, we estimate a time accelerated hazard model with a log-normal parametric specification [@box-steffensmeier_durationmodelsproportional_2001; @box-steffensmeier_eventhistorymodeling_2004].^[Time accelerated hazard models have been used to estimate factors associated with the duration of interstate or intrastate conflict by @slantchev_howinitiatorsend_2004, @martinezmachain_aircampaignduration_2015, @caverley_militarytechnologyduration_2017, and @chiba_militarycoalitionscrisis_2019, among others. Although other scholars have estimated a Cox proportional hazard model [@allen_timebombsestimating_2007; @beardsley_winningbomb_2009; @fortna_terroristswinrebels_2015], a proportional hazard assumption test and the Schoenfeld residuals (provided in the appendix) show a Cox proportional hazard model is inappropriate here [@box-steffensmeier_nonproportionalhazardsevent_2003]. Other parametric specifications for the hazard model are estimated in the appendix. The log-normal is most appropriate as it has the smallest AIC.]

```{r dv, fig.cap = "Distribution of dependent variables", results = 'asis', fig.show = "hold", out.width = "80%"}
df_crisis %>%
  dplyr::mutate(intensity = dplyr::recode_factor(sevviosy,
                                                 "1" = "No violence",
                                                 "2" = "Minor clash",
                                                 "3" = "Serious clash",
                                                 "4" = "Full-scale war")) %>%
  ggplot(aes(x = intensity)) +
  geom_bar() +
  labs(title = "Dependent Variable: Severity of Violence",
       x = "Severity of Violence",
       y = "Number of crises"
       ) +
  theme_minimal() +
  theme(text = element_text(size = 14))

df_crisis %>%
  ggplot(aes(x = brexit)) +
  geom_histogram() +
  labs(title = "Dependent Variable: Crisis Duration",
       x = "Crisis Duration (Days)",
       y = "Number of crises"
       ) +
  scale_x_continuous(breaks = seq(00, 1500, by = 250)) +
  theme_minimal() +
  theme(text = element_text(size = 14))
```

One shortfall of using the crisis as the unit of analysis is that many conventional control variables that exist at the actor or sub-state level cannot be appropriately measured given difficulty in theoretically motivating methods of aggregating those measures to the crisis-level [@petersen_multipartydisputesprobability_2004, 91].^[This includes state-level variables like regime type and dyad-level variables like rivalry. There is a vast and well-developed literature on micro-foundations associated with the intensity and duration of conflict like leader traits and psychology. See, among others, @mcdermott_prospecttheorypolitical_2004, @saunders_transformativechoicesleaders_2009, @chiozza_leadersinternationalconflict_2011, @horowitz_howpriormilitary_2014, @yarhi-milo_armallypatron_2016, @renshon_emotionsmicrofoundationscommitment_2017, @stein_microfoundationsinternationalrelations_2017, and @yarhi-milo_whofightsreputation_2018. We bracket discussion of the relationship between individual attributes and the choice of military domains for future research.] Nonetheless, both models include a battery of controls that are theoretically motivated giving existing literature concerning factors that influence the intensity and duration of international crises. We include a control for the number of crisis actors, since more participants in an international crisis may make it more difficult to get a mutually acceptable agreement, thus increasing the intensity and duration of a crisis [@petersen_multipartydisputesprobability_2004]. We include a control for whether or not a crisis is part of a protracted conflict, which prior research has hypothesized produces more violent crises since they are part of a process that is more difficult to resolve [@azar_protractedsocialconflict_1978; @brecher_studycrisis_2000].^[ICB defines a protracted conflict as "an environment of ongoing disputes among adversaries, with fluctuating interaction ranging from violence to near-tranquility, multiple issues and spillover effects on all aspects of their relations, and the absence of mutually-recognized or anticipated termination (the Arab-Israeli conflict, 1947-)" [@brecher_internationalcrisisbehavior_2020].] We include a control for whether the value that a crisis actor felt was threatened was territorial in nature since territorial conflicts like border wars are more violent and difficult to resolve [@vasquez_whyneighborsfight_1995; @vasquez_territorialdisputesprobability_2001; @owsiak_clearinghurdleborder_2013]. Territorial conflicts could also involve more similar capabilities by opposing belligerents since land and air forces may be most relevant in holding or taking territory. We include a control for whether the crisis was motivated by ethnic differences since ethnic conflicts are more violent and difficult to resolve [@ben-yehuda_ethnicactorsinternational_2006; @mishali-ram_ethnicdiversityissues_2006]. We include a control for the the power disparity between both sides of a crisis, which is a composite measure of population, GNP, major power alliances, territorial size, military capability, and alliance capability [@quinn_powerplaymediation_2006]. We include a control for whether one of the two superpowers, the United States or Soviet Union, was involved in the crisis [@colaresi_strategicrivalriesprotracted_2002]. Lastly, we control for contiguity which identifies whether or not the primary crisis actors share a border [@bremer_dangerousdyadsconditions_1992; @vasquez_distinguishingrivalsthat_1996]. Figure 4 shows the bivariate correlation between each of the two dependent variables (violence severity and crisis duration) and the independent variable of interest (cross-domainness).

## Results
The full model results are shown in Table \@ref(tab:model). For each dependent variable we first estimate a bivariate model containing only crossdomainness as an independent variable. We then conduct full models that include all control variables. For the first two models concerning crisis intensity, the results show that cross-domainness is negatively associated with intensity of violence with statistical significance of at least the 0.05 level. As the coefficients of ordered logistic regression are difficult to interpret, the odds ratio coefficient for the crossdomainness variable (0.66) indicates that the odds of a crisis with a serious clash or full scale war are 44% lower than the odds of experiencing violence with a minor clash or no violence if the actors use completely dissimilar means.^[A complete table of the odds ratios is provided in the appendix]. These results provide evidence consistent with theories of cross-domain deterrence rather than cross-domain escalation --- crisis in which opponents take military actions in dissimilar domains are less violent than those in which opponents take military actions in the same domains. As models 3 and 4 are parametric time accelerated hazard models, the coefficients detail the likelihood of a crisis ending on a given day. Positive coeffients indicate a variable is associated with a longer crisis while negative coefficients mean the variable is associated with a shorter crisis. The coefficients for cross-domainness are negative, but not statistically significant at the 0.05 level once control variables are added to the model, meaning we cannot reject the null hypothesis that cross-domainness has no discernable effect on the duration of a crisis. The only statistically significant coefficients for crisis duration are for the number of crisis actors and ethnic conflict. Both are statistically significant at the 0.05 level and positive, meaning that --- consistent with prior research --- crises with more actors and that are ethnic disputes are longer than crises with fewer actors or over non-ethnic issues.

```{r, model, results = 'asis'}
# model 1 bivariate severity
df_m1 <- df_crisis %>%
  dplyr::select(sevviosy, crossdomain_jaccard) %>%
  dplyr::mutate(sevviosy = as.factor(sevviosy))

m1 <- polr(sevviosy ~ .,
           data = df_m1,
           method = "probit",
           Hess = TRUE)

# model 2 full severity
df_m2 <- df_crisis %>%
  dplyr::select(sevviosy, crossdomain_jaccard, noactr, powdissy, protrac, territ, majpwrinv, ethnic, contig) %>%
  dplyr::mutate(sevviosy = as.factor(sevviosy))

m2 <- polr(sevviosy ~ .,
           data = df_m2,
           method = "probit",
           Hess = TRUE)

# model 3 bivariate
df_m3 <- df_crisis %>%
  dplyr::select(brexit, crossdomain_jaccard)

m3 <- survival::survreg(survival::Surv(brexit) ~ .,
                      dist = 'lognormal',
                      data = df_m3)

# model 4 full duration
df_m4 <- df_crisis %>%
  dplyr::select(brexit, crossdomain_jaccard, noactr, powdissy, protrac, territ, majpwrinv, ethnic, contig)

m4 <- survival::survreg(survival::Surv(brexit) ~ .,
                        dist = 'lognormal',
                        data = df_m4)

# Table of results
models <- list(m1, m2, m3, m4)

texreg::texreg(models,
               stars = c(0.01, 0.05, 0.1),
               omit.coef = "(1|2|3|4)",
               custom.coef.names = c("Cross-domainness",
                                     "No. of actors",
                                     "Power Dissimilarity",
                                     "Protracted Crisis",
                                     "Territorial Crisis",
                                     "Major Power Involv.",
                                     "Ethnic Crisis",
                                     "Contiguity",
                                     "Intercept",
                                     "Log (scale)"),               
               custom.header = list("Violence Intensity" = 1:2, "Crisis Duration" = 3:4),
               custom.note = ("\\parbox{.75\\linewidth}{\\vspace{2pt}%stars. \\\\ 
                              Models 1 and 2 are ordered probit models and models 3 and 4 are log-normal accelerated failure time models.}"),
               float.pos = "h!",
               label="tab:model",
               use.packages = FALSE)
```

These findings are consistent across a range of alternate modeling decisions detailed in the appendix. We run additional models using a binary dependent variable for violence severity, a modeling decision used in prior work on ICB crisis violence [@hewitt_democraciesinternationalcrisis_1996; @chiozza_internationalconflicttenure_2004]. Despite the appropriateness of ordered probit and a log-normal parametric specification for the duration model, we also find consistent results using ordered logistic and OLS regression for the severity of violence variable and using other parametric specifications for the duration dependent variable.

Crises in which belligerents interact in dissimilar military domains are less violent, but neither longer nor shorter, that crises in which belligerents interact with like-means. While contemporary cross-domain conflicts with new modes of conflicts have ignited pessimism about the escalatory potential of new modes of conflict, the empirical evidence should give interested parties some confidence that they can respond to their adversaries' "apples" with their own "oranges" without that being associated with a bloodier and longer contest. As this is a large-n study using observational data, it is difficult to determine the causal direction of the relationship. This paper has proposed one mechanism by which cross-domain interactions could intensity the severity and duration of a crisis. The causal arrow may of course run the other way. It could be that the more a state cares about the outcome of the crisis, the more likely they are to bring their best military assets to the fight. But doing so may cause more blood to be spilled since it involving bringing mobility, stealth, and complexity to the battlefield.

# Implications of bringing a knife to a gun fight
Emerging interest in understanding cross-domain military interactions is well-deserved, given the frequency with which these interactions occur. But this research need not be spurned by, not limited to, the study of new domains made possible because of emerging technologies. Crises have always been cross-domain and theories that concern interactions involving new domains of conflict need not reinvent the wheel since cross-domain interactions have been a consistent regularity for the past century. Different ways of defining and thinking about domains may produce new insights and certainly new domains should be compared to more traditional ones, but the findings here represent an important and thus far underrepresented contribution to that discussion.

Concerns about the escalation dynamic of cyber-kinetic interactions are loosely based off of theories of deterrence and spiral models of conflict that have rarely accounted for the strategic interaction of the military domains used during conflict. The common assumption held by pessimists---that cross-domain interactions risk a dangerous spiral because of potential miscommunication over proportionality, stake, and resolve. Rather, cross-domain interactions contain elements of the stability-instability paradox, where one side's willingness to operate in a new domain of conflict---a potential indicator of a willingness to escalate---creates conditions that lessen observed crisis escalation. This paper builds on the state of the art by furthering our understanding of conflict escalation. At least one part of the story in explaining why some conflicts spiral while others deter concerns the interaction between the domains in which each side's military is taking action. These insights help us answer the question about whether emerging technologies will make conflicts worse or whether they will be stabilizing [@sechser_emergingtechnologiesstrategic_2019; @talmadge_emergingtechnologyintrawar_2019]. Attempts to use full-spectrum combined arms forces may make sense for a strategic standpoint if they increase the probability of victory---an outcome not examined here--- but those doing so should consider preparing for a bloodier war.

The escalatory effects of the tools used in a crisis are fundamental to understanding the evolution of conflict in the 21\textsuperscript{st} century. These findings have implications for a number of contemporary policy debates. Recent work has investigated how other states can best combat Chinese naval expansion [@beckley_emergingmilitarybalance_2017; @talmadge_emergingtechnologyintrawar_2019; @cunningham_maritimerungescalation_2020]. As the United States considers anti-access/anti-denial (A2/AD) in East Asia vis-a-vis competition with China, strategies that seek to take advantage of full-spectrum combined-arms forces where China does not, blending multiple military tools together as done with AirSea Battle, may invoke a less bloody contest that matching China's moves. As NATO considers how to respond to Russian gray zone aggression, whether "little green men" in Crimea or cyber attacks against Estonia, responding with other non-cyber or non-special operations capabilities may not be as escalatory as alarmists argued after Israel responded to Hamas cyberattacks with missile strikes.

There is much more to be said about how states fight, both as an independent variable explaining other outcomes of interest and also as a dependent variable of interest itself. As the primary contribution of this paper concerns new data on domain-interactions over the past century, we hope part of its value comes in opening up many other avenues for research concerning questions this paper raises but does not answer. Not all cross-domain interactions are created equal. While the theories tested here concern general cross-domain escalation dynamics, a land vs air conflict may turn out differently than a land vs sea conflict, although both are considered "cross-domain". There are specific contexts where that may be especially true, like the India Pakistan context where Pakistani concerns about Indian mobilization of integrated battle groups may shift the cross-domain dynamic in ways not explained here. Future work should examine the sequence of when actors take actions in different domains, since that matters for theories of first-mover advantage and unintended escalation [@quester_offensedefenseinternational_1977; @reiter_howwarsend_2009; @garfinkel_howdoesoffensedefense_2019]. The data as presented here makes simplifying aggregations regarding the domains in which states take actions during international crises. Future work can disaggregate each of those. By looking at military units instead of domains, theories can be further refined since, for example, not all naval capabilities are the same in terms of things like signaling or warfighting. Actions can also be further disaggregated, as further work with the underlying data could parse out and evaluate differences in types of actions like mobilizations, raises in alert level, bombardments, and occupations. The theory and findings here are also limited to military domains, but the interaction between a state's use of military, economic, and diplomatic tools would also contribute to our collective understanding of crisis escalation and duration. Lastly, this paper investigates military domains as an independent variable in the hopes of furthering our understanding of how the choice of military domains is associated with crisis intensity and duration. We have not explained why states choose to operate in some military domains as opposed to others. Future work should think of the military domains or crisis behavior as a dependent variable and use this novel data to explain why states choose the operate in the domains they do.

# Conclusion
Epidemiologists spend much of their time identifying factors that increase the risk of particular health problems in individuals. An important component of their work relies on the proper identification of events and conditions that serve as an indicator that health problems will soon follow. Similarly, scholars of international relations are interested in identifying factors that indicate that the outbreak of conflict is increasingly likely [@senese_stepswarempirical_2008]. The goal in this case is not only to causally identify factors that make war more likely so that we can better understand events of the past, but also to improve our ability to forecast the likelihood of conflict in the future [@valeriano_pathwaysinterstatewar_2010]. The military domains in which states contest one another and the similarity of those domains constitutes an important piece of that puzzle. Cross-domain contests in which states engage each other with dissimilar military means are less violent and no longer than crises in which states respond to each other in-kind. There may be other reasons to sound the alarm over emerging technologies, but the historical record shows that responding to aggression in a new and unfamiliar domain in itself does not constitute a higher risk of conflict. Cross-domain interactions are more associate with deterring conflict than enflaming it and, contrary to popular wisdom, always have been.

# References
